---
title: "Tuesday Night Football"
subtitle: "The Annual Analytics Review"
author: "TNF Analytics Platform"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: true
    latex_engine: pdflatex
    fig_caption: false
header-includes:
  - \usepackage{tcolorbox}
  - \usepackage{booktabs}
  - \usepackage{float}
  - \newtcolorbox{insight}{colback=orange!8,colframe=orange!70!black,boxrule=0.8pt,arc=2mm,left=2mm,right=2mm,top=1mm,bottom=1mm}
  - \floatplacement{figure}{H}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.width = 9, fig.height = 5.6, dpi = 150,
                      fig.pos = "H", out.width = "100%")
suppressPackageStartupMessages({
  library(tidyverse); library(tidygraph); library(ggraph); library(kableExtra)
})
root <- normalizePath("..")
for (f in list.files(file.path(root, "R"), full.names = TRUE)) source(f)
db <- build_database(file.path(root, "data-raw/team_sheets_raw.txt"),
                     file.path(root, "data-raw/player_aliases.csv"))
ms <- match_summary(db)
tr <- team_record(db)
ps <- player_stats(db) |> filter(!is_guest)

insight <- function(...) {
  txt <- paste0(...)
  txt <- gsub("%", "\\\\%", txt)   # % starts a comment in raw LaTeX
  txt <- gsub("&", "\\\\&", txt)
  cat("```{=latex}\n\\begin{insight}\n\\textbf{Insight.} ", txt,
      "\n\\end{insight}\n```\n")
}
tk <- function(d, ...) {
  kbl(d, booktabs = TRUE, linesep = "", ...) |>
    kable_styling(latex_options = c("striped", "HOLD_position"),
                  font_size = 9)
}
```

# Executive Summary

Tuesday Night Football has now completed **`r ms$total_matches` matches**
between `r format(ms$first_match, "%d %B %Y")` and
`r format(ms$last_match, "%d %B %Y")`, with **`r ms$total_players` different
players** pulling on (or refusing to pull on) a bib. Between them they have
scored **`r ms$total_goals` goals** — an average of
`r round(ms$goals_per_match, 1)` per night — in front of an average
"crowd" of `r round(ms$mean_attendance, 1)` participating players.

```{r exec-facts, results='asis'}
facts <- generate_discoveries(db)
cat(paste0("- ", facts), sep = "\n")
```

# Season Overview

```{r results-timeline}
plot_results_timeline(db)
```

*Each bar is one Tuesday. Bars pointing up are Bibs wins, bars pointing down
are Non-Bibs wins, and the bar's height shows how comfortable the win was.
Grey dots on the line are draws.*

```{r alluvial, fig.height=5}
plot_results_alluvial(db)
```

*This flow diagram takes each side's matches (left) and shows how they ended
up (right). The thicker the ribbon, the more matches took that path.*

```{r season-callout, results='asis'}
bibs <- tr[tr$team == "Bibs", ]
insight("Bibs lead the all-time series: ", bibs$wins, " wins to ",
        tr$wins[tr$team == "Non-Bibs"], " with ", bibs$draws,
        " draws. Their ", fmt_pct(bibs$win_pct),
        " win rate carries a 95% confidence interval of ",
        fmt_pct(bibs$lower), "--", fmt_pct(bibs$upper),
        ", so the jury is technically still out.")
```

# Match Statistics

## The record so far

```{r team-record}
tr |>
  transmute(Team = team, P = played, W = wins, D = draws, L = losses,
            GF = goals_for, GA = goals_against, GD = goal_diff,
            `Win %` = fmt_pct(win_pct),
            `95% CI` = paste0(fmt_pct(lower), "--", fmt_pct(upper))) |>
  tk(caption = "All-time record by side")
```

```{r team-streaks}
team_streaks(db) |>
  transmute(Team = team, `Longest win streak` = longest_win_streak,
            `Longest unbeaten` = longest_unbeaten,
            `Longest losing run` = longest_losing_streak,
            `Current run` = current) |>
  tk(caption = "Streaks by side")
```

## Biggest victories

```{r biggest}
biggest_victories(db) |>
  transmute(Date = format(date, "%d %b %Y"), Score = score, Winner = winner,
            Margin = margin, Attendance = attendance) |>
  tk(caption = "Most emphatic wins (score shown as Bibs--Non-Bibs)")
```

## Scoring

```{r goals-violin, fig.height=5}
plot_goals_violin(db)
```

*Each dot is one team's goal tally in one match. The coloured shapes show
where tallies bunch up: both sides usually score between three and nine, with
the occasional night of total madness.*

```{r goals-ridge, fig.height=6}
plot_goals_ridgeline(db)
```

*Each ridge is a month. The further right the bump, the more goals flew in
that month — the spike on the far right is the legendary 18--19 night of
25 June.*

## Attendance

```{r attendance}
plot_attendance_over_time(db)
```

*Each dot is the number of players who turned up that Tuesday. The dashed
line is the season average, so dots above it were busy nights.*

```{r format, fig.height=4.5}
plot_format_over_time(db)
```

*Each dot is a match night, categorised by format: with two teams of roughly
equal size, the night is either 5-, 7- or 8-a-side.*

# Player Statistics

## The headline table

```{r player-table}
ps |>
  arrange(desc(appearances)) |>
  transmute(Player = player_name, Apps = appearances, W = wins, D = draws,
            L = losses, `Win %` = fmt_pct(win_pct), GF = goals_for,
            GA = goals_against, GD = goal_diff,
            Streak = current_streak,
            `Small?` = ifelse(small_sample, "*", "")) |>
  tk(caption = "All players: appearances, results and goals while on the pitch. * = fewer than 8 appearances, treat with caution") |>
  column_spec(1, bold = TRUE)
```

```{r apps-lollipop, fig.height=7}
plot_appearances_lollipop(db)
```

*The club's most devoted attendees. Each stick shows the total number of
Tuesdays a player has turned out for.*

## Win rates (and how sure we can be)

```{r winpct, fig.height=7}
plot_win_pct_ci(db)
```

*The dot is each player's win rate; bigger dots mean more matches. The
whiskers show the range where the "true" win rate plausibly lives — the wider
the whisker, the less the sample tells us. Almost every whisker crosses the
50% line, which is football's way of saying it is nicely balanced.*

```{r goal-diff, fig.height=6.5}
plot_goal_diff(db)
```

*Green bars: the team outscores opponents when this player plays. Red bars:
the opposite. This blends personal quality, teammates and pure luck — which
is exactly why it makes for good arguments.*

```{r streaks-fig, fig.height=5.5}
plot_streaks(db)
```

*The longest unbroken run of wins each player has ever been part of.*

```{r player-callout, results='asis'}
best <- ps |> filter(appearances >= 8) |> slice_max(win_pct, n = 1, with_ties = FALSE)
loyal <- ps |> filter(appearances >= 8) |> slice_max(attendance_consistency, n = 1, with_ties = FALSE)
insight(best$player_name, " tops the win table among regulars at ",
        fmt_pct(best$win_pct), ", while ", loyal$player_name,
        " is the metronome of the group, present for ",
        fmt_pct(loyal$attendance_consistency),
        " of all matches since debuting.")
```

# Partnerships

```{r partnerships-fig, fig.height=6}
plot_partnerships(db)
```

*The best-performing duos among pairs who have shared a team at least five
times. The label shows wins out of matches together; whiskers again show the
uncertainty.*

```{r pairs-table}
pair_stats(db, 8) |>
  arrange(desc(together)) |>
  head(12) |>
  transmute(Pair = paste(player_a, "+", player_b), Together = together,
            W = wins, D = draws, L = losses, `Win %` = fmt_pct(win_pct),
            GD = goal_diff) |>
  tk(caption = "Most frequent partnerships (8+ matches together)")
```

```{r co-occ, fig.height=8.5, fig.width=9.5}
plot_co_occurrence(db)
```

*How often every pair of regulars has been on the same team. Deep orange
squares are near-inseparable teammates; blank squares are pairs the bib
randomiser has kept apart.*

# Team Chemistry

```{r combos-trio}
combo_stats(db, 3, 4) |>
  head(10) |>
  transmute(Trio = combo, Together = together, W = wins, D = draws, L = losses,
            `Win %` = fmt_pct(win_pct)) |>
  tk(caption = "Best trios (4+ matches together)")
```

```{r combos-quartet}
combo_stats(db, 4, 3) |>
  head(8) |>
  transmute(Quartet = combo, Together = together, W = wins,
            `Win %` = fmt_pct(win_pct)) |>
  tk(caption = "Best quartets (3+ matches together)")
```

```{r combos-quintet}
combo_stats(db, 5, 3) |>
  head(6) |>
  transmute(Quintet = combo, Together = together, W = wins,
            `Win %` = fmt_pct(win_pct)) |>
  tk(caption = "Best quintets (3+ matches together)")
```

```{r stability, results='asis'}
stab <- lineup_stability(db)
insight("Squad turnover is real: on average only ",
        fmt_pct(mean(stab$jaccard)),
        " of one Tuesday's squad overlaps with the next (Jaccard similarity), ",
        "so the bib randomiser gets fresh material every week.")
```

## Who beats whom

```{r h2h-fig, fig.height=8, fig.width=9.5}
plot_h2h_heatmap(db)
```

*Read along a row: the colour shows how often that player's team beats the
column player's team when they are on opposite sides. Orange = dominance,
blue = suffering. Only match-ups with four or more meetings are shown.*

```{r h2h-sentences, results='asis'}
cat(paste0("- ", h2h_sentences(db, min_meetings = 6, top = 8)), sep = "\n")
```

# Network Analysis

```{r network, fig.height=8, fig.width=10.5}
plot_network(db)
```

*Every line connects two players who have shared a team at least three times.
Thick lines are long-running partnerships; green lines win together, red
lines lose together. The node colours are "communities" — clusters of players
the algorithm thinks belong to the same social orbit.*

```{r centrality-fig, fig.height=5.5}
plot_centrality(db)
```

*PageRank — the algorithm Google made famous — applied to Tuesday night:
players score highly for repeatedly sharing teams with other well-connected
players.*

```{r centrality-table}
network_metrics(db) |>
  head(12) |>
  transmute(Player = name, Degree = degree, Strength = strength,
            Betweenness = round(betweenness, 1),
            Closeness = round(closeness, 3),
            PageRank = round(pagerank, 4)) |>
  tk(caption = "Network centrality (top 12 by PageRank)")
```

# Time Trends

```{r rolling}
plot_rolling_winpct(db)
```

*Form over time: each line shows the share of the last five matches each side
has won. When the orange line is on top, Bibs are the team to be on.*

```{r dominance}
plot_bib_dominance(db)
```

*The season-long tug of war: Bibs' share of all available points so far.
Above the dashed line, Bibs lead the season; below it, Non-Bibs do.*

```{r cumulative, fig.height=6.5}
plot_cumulative_appearances(db)
```

*Every player's appearance count as it grows through the season. The steepest
lines belong to the ever-presents.*

```{r bump, fig.height=6}
plot_monthly_bump(db)
```

*Month by month, who turned up most? Rank 1 means the most appearances that
month among the club's eight busiest players.*

```{r timeline, fig.height=7.5}
plot_player_timeline(db)
```

*The whole season in one picture: each row is a player (newest at the top),
each dot an appearance, coloured by how the night ended.*

```{r retention, results='asis'}
ret <- player_retention(db)
newc <- newcomer_rate(db)
insight("Month-to-month retention averages ", fmt_pct(mean(ret$retention)),
        ": most of the squad comes back for more. Meanwhile ",
        sum(newc$debutants[-1]),
        " new players debuted after the opening month, keeping the pool fresh.")
```

# Interesting Facts

```{r facts, results='asis'}
cat(paste0(seq_along(facts), ". ", facts), sep = "\n")
```

# Appendix

## Full results

```{r all-results}
match_results_table(db) |>
  transmute(Date = format(date, "%d %b %Y"), Score = score, Winner = winner,
            Attendance = attendance, Note = coalesce(note, "")) |>
  tk(caption = "Every match (score shown as Bibs--Non-Bibs)")
```

## Notable events

```{r events}
db$events |>
  left_join(db$players, by = "player_id") |>
  transmute(Date = format(date, "%d %b %Y"), Event = event_type,
            Player = coalesce(player_name, "—"), Detail = detail) |>
  tk(caption = "Recorded events")
```

## Player alias table

```{r aliases}
db$player_aliases |>
  transmute(Alias = alias, `Canonical name` = canonical,
            `Ambiguous?` = ifelse(ambiguous, "yes", ""),
            Note = coalesce(note, "")) |>
  tk(caption = "Alias lookup used to standardise names")
```

## Data notes and assumptions

- The number in brackets after each team name is that team's **goals**, not
  its player count (attendances and squad sizes are derived from the listed
  names).
- Dates are day/month/year; two-digit years are assumed to be 20xx.
- Ambiguous first names (e.g. "Tom" on 27 January) are resolved by
  elimination within the match; every automatic decision is logged in
  `data/parse_issues.csv`.
- "Rik" is treated as a typo for Riz Khan; "Comical A-Lee" as a nickname for
  Lee; "Lee's mate" as an unnamed guest.
- Win percentages come with 95% Wilson score intervals. With `r ms$total_matches`
  matches, almost all player-level differences are within the noise — which
  has never stopped a good argument.

```{r issues}
db$parse_issues |>
  transmute(Date = format(date, "%d %b %Y"), Severity = severity,
            Message = message) |>
  tk(caption = "Parser and alias-resolution log")
```
